---
title: "Gather Results"
author: "Dana Melamed"
date: "`r Sys.Date()`"
output: html_document
---

Checkout modelling results for the entire growing season period of 2012.
Then, apply the model to other years, and calculate a 5 year average.

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.align = 'center')
knitr::opts_chunk$set(fig.width = 9)
knitr::opts_chunk$set(fig.height = 7)
library(swnsmodelr)
par(mfrow=c(2,2))
```

## Model temperature with each input variable individually

```{r}
# Prepare the data
# Make a list of input rasters
in_folder <- 'f:\\data\\rasters\\input_rasters\\'
raster_names <- list('dem700m','cp700m','east700m','north700m','tpi5000m700m',
                     'asp700m','slp700m')
ext = '.tif'
rasters_list <- lapply(X=paste0(in_folder,raster_names,ext),FUN=raster::raster)

#transform
rasters_list[[2]]<-rasters_list[[2]]^0.5
names(rasters_list[[2]]) <- 'cp700m'
rasters_list[[6]] <- rasters_list[[7]]*cos(rasters_list[[6]])
names(rasters_list[[6]]) <- 'asp700m'

# brick rasters
rasters_brick2 <- raster::brick(rasters_list)

# extract rasters to dataframe
stations_df <- swns_stations_df %>% dplyr::select(date_time,
                                           stationid,
                                           EASTING,
                                           NORTHING,
                                           temp_min,
                                           temp_max)
df_in <- stations_df %>% 
  dplyr::select(stationid,temp_min,temp_max,EASTING,NORTHING,date_time) %>%
  extract_constant_raster_values(rasters_list) %>% add_date_columns()
```





Create dataframe for each year
```{r}

# calculate gdd10 in dataframe
df <- df_in %>%
  mutate(temp_mean=(temp_min+temp_max)/2)%>%
  dplyr::filter(stationid %in% nscc_stations_list |
                  stationid == '47187' |
                  stationid == '6354')%>%
  dplyr::filter(stationid!='S100') %>% # 2012
  dplyr::filter(stationid!='S160') %>% # 2012, early in season/cold
  dplyr::filter(stationid!='S20')  %>% #2012
  dplyr::filter(stationid != 'S60') %>% #2012
  dplyr::filter(stationid != 'S80') %>% #2012
  dplyr::filter(stationid != 'CL1') %>%
  mutate(gdd10_daily=ifelse(temp_mean>10,temp_mean-10,0)) %>%
  group_by(year,stationid) %>%
  mutate(gdd10 = sum(gdd10_daily, na.rm = TRUE))

df_12 <- df %>% dplyr::filter(year=='2012'&between(month,4,11)) %>%
  group_by(stationid)%>%
  dplyr::filter(date_time == max(date_time))

df_13 <- df %>% dplyr::filter(year=='2013'&between(month,4,11)) %>%
  group_by(stationid)%>%
  dplyr::filter(date_time == max(date_time))

df_14 <- df %>% dplyr::filter(year=='2014'&between(month,4,11)) %>%
  group_by(stationid)%>%
  dplyr::filter(date_time == max(date_time))

df_15 <- df %>% dplyr::filter(year=='2015'&between(month,4,11)) %>%
  group_by(stationid)%>%
  dplyr::filter(date_time == max(date_time))

df_16 <- df %>% dplyr::filter(year=='2016'&between(month,4,11)) %>%
  group_by(stationid)%>%
  dplyr::filter(date_time == max(date_time))
  
```

### Position: East, North, (East,North), and Coastal Proximity

```{r}
m_list <- list()
m_list[[1]]<- gam(gdd10~s(east700m),data=df_12,method='REML')
m_list[[2]]<- gam(gdd10~s(north700m),data=df_12,method='REML')
m_list[[3]]<- gam(gdd10~s(east700m)+s(north700m),data=df_12,method='REML')
m_list[[4]]<- gam(gdd10~s(east700m,north700m),data=df_12,method='REML')
m_list[[5]]<- gam(gdd10~s(cp700m),data=df_12)
m_list[[6]]<- gam(gdd10~s(east700m,north700m,k=10),data=df_12,method='REML')

for(m in m_list){
  
  gam_raster <- raster::predict(rasters_brick2,m)
  plot(gam_raster, main=m$formula, col=rev(rainbow(100)[1:70]))
  points(swns_stations_sp)
  par(mfrow=c(2,2),mar=c(5,6,2,2)+0.1)
  plot.gam(m, residuals = TRUE, pch = 1, cex = 1, shift = coef(m)[1])
  par(mfrow=c(1,1))
}

```

### Land Surface: DEM, Slope, Aspect (Guan et al., 2013), TPI5000m700m

```{r}
m_list <- list()
m_list[[1]]<- gam(gdd10~s(dem700m),data=df_12,method='REML')
m_list[[2]]<- gam(gdd10~s(asp700m),data=df_12,method='REML')
m_list[[3]]<- gam(gdd10~s(slp700m),data=df_12,method='REML')
m_list[[4]]<- gam(gdd10~s(tpi5000m700m),data=df_12,method='REML')
for(m in m_list){
  gam_raster <- raster::predict(rasters_brick2,m)
  plot(gam_raster, main=m$formula, col=rev(rainbow(100)[1:70]))
  par(mfrow=c(2,2))
  plot.gam(m, residuals = TRUE, pch = 1, cex = 1, shift = coef(m)[1])
  par(mfrow=c(1,1))
}
```

## Combine Variables

Above, it seems the reasonable rasters to model with are (East,North), Coastal Proximity, DEM

### (East,North) + (x)

```{r}
m_list <- list()
m_list[[1]] <- gam(gdd10~s(east700m,north700m) +s(cp700m),data=df_12,method='REML')
m_list[[2]] <- gam(gdd10~s(east700m,north700m)+ s(dem700m),data=df_12,method='REML')

for(m in m_list){
  
  gam_raster <- raster::predict(rasters_brick2,m)
  plot(gam_raster, main=m$formula, col=rev(rainbow(100)[1:70]))
  par(mfrow=c(2,2))
  plot.gam(m, residuals = TRUE, pch = 1, cex = 1, shift = coef(m)[1])
  par(mfrow=c(1,1))
}
```

### te(DEM, cp) \| s(dem) + s(cp)

```{r}
m_list <- list()
m_list[[1]] <- gam(gdd10~s(dem700m)+s(cp700m),data=df_12,method='REML')
m_list[[2]] <- gam(gdd10~te(dem700m,cp700m),data=df_12,method='REML')

for(m in m_list){
  gam_raster <- raster::predict(rasters_brick2,m)
  plot(gam_raster, main=m$formula, col=rev(rainbow(100)[1:70]))
  par(mfrow=c(2,2))
  plot.gam(m, residuals = TRUE,     pch = 1, cex = 1, shift = coef(m)[1])
  par(mfrow=c(1,1))
}
```

### Combine... s(east,north) + s(dem,cp)|s(dem) +s(cp)

```{r}
m_list <- list()
m_list[[1]] <- gam(gdd10~te(dem700m,cp700m)+s(east700m,north700m),data=df_12,method='REML')
m_list[[2]] <- gam(gdd10~s(dem700m)+s(cp700m)+s(east700m,north700m),data=df_12,method='REML')

for(m in m_list){
  
  gam_raster <- raster::predict(rasters_brick2,m)
  plot(gam_raster, main=m$formula, col=rev(rainbow(100)[1:70]))
  par(mfrow=c(2,2))
  plot.gam(m, residuals = TRUE,  pch = 1, cex = 1, shift = coef(m)[1])
  par(mfrow=c(1,1))
}
```



### Knots
Define knots so that wiggliness is increased
```{r}
m_list <- list()
m_list[[1]] <- gam(gdd10~te(dem700m,cp700m,k=5)+s(east700m,north700m,k=10),data=df_12,method='REML')
m_list[[2]] <- gam(gdd10~s(dem700m,k=30,bs='cr')+s(cp700m,k=5,bs='cr')+s(east700m,north700m,k=20),data=df_12,method='REML',family=gaussian())

m<-gam(gdd10~s(cp700m),data=df_12)
for(m in m_list){
  
  gam_raster <- raster::predict(rasters_brick2,m)
  plot(gam_raster, main=m$formula, col=rev(rainbow(100)[1:70]))
  par(mfrow=c(2,2))
  plot.gam(m, residuals = TRUE, pch = 1, cex = 1, shift = coef(m)[1])
  par(mfrow=c(1,1))
}
```

Other years
```{r}
m_list <- list()
m_list[[1]] <- gam(gdd10~s(dem700m,k=30,bs='cr')+s(cp700m,k=5,bs='cr')+s(east700m,north700m,k=20),data=df_13,method='REML',family=gaussian())

m_list[[2]] <- gam(gdd10~s(dem700m,k=30,bs='cr')+s(cp700m,k=5,bs='cr')+s(east700m,north700m,k=20),data=df_14,method='REML',family=gaussian())

m_list[[3]] <- gam(gdd10~s(dem700m,k=30,bs='cr')+s(cp700m,k=5,bs='cr')+s(east700m,north700m,k=20),data=df_15,method='REML',family=gaussian())

m_list[[4]] <- gam(gdd10~s(dem700m,k=30,bs='cr')+s(cp700m,k=5,bs='cr')+s(east700m,north700m,k=20),data=df_16,method='REML',family=gaussian())


for(m in m_list){
  
  gam_raster <- raster::predict(rasters_brick2,m)
  plot(gam_raster, main=m$formula, col=rev(rainbow(100)[1:70]))
  points(swns_stations_sp)
  par(mfrow=c(2,2))
  plot.gam(m, residuals = TRUE, pch = 1, cex = 1, shift = coef(m)[1])
  par(mfrow=c(1,1))
}
```