---
title: "swnsmodelr R Package Manual"
author: "Dana Melamed"
date: "June 7, 2018"
output:
  rmarkdown::word_document:
    fig_caption: yes
indent: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(swnsmodelr)
library(ggplot2)
```

## Introduction

  The swnsmodelr package was designed for users to easily model SWNS weather station data with rasters and models of their chosing. The package contains functions and weather station data, and input rasters are contained within the package's R project. 
This document follows a workflow of preparing data, generating models, and generating outputs. The objects generated at the various steps in the example are stored in swnsmodelr. 



## Exract Rasters to Weather Station Data

  The first step in modelling weather station data with rasters is to extract raster values to corresponding weather station records.  For many rasters, (ex: elevation, proximity to coast, aspect) the value will be same every day for a station. Other rasters (ex: solar radiation) change on a temporal basis, and will have a unique value for each station every day. The two type of rasters, constant, and temporally changing, will be treated differently in the value extract process.

### Weather Stations Data Frame
  The data frame, `swns_stations_df`, stored within swnsmodelr, contains SWNS daily temperature data from 2012 - 2017. The data frame has columns for: station I.D. (`stationid`), date (`date_time`), minimum temperature (`temp_min`), maximum temperature (`temp_max`), and mean temperature (`temp_mean`). One row of data represents one daily record for one station. Use `str()`, to see the column names and variables types. 
  
```{r message=FALSE, warning=FALSE}
str(swns_stations_df)
```


### Constant Rasters

  The constant rasters should be stored as a list of raster objects. The code below is an example of how to create a list of raster objects from rasters available in the package project. The rasters are from the `200` subdirectory within the `Rasters` folder. In this subdirectory, rasters have a spatial resolution of 200m.  
```{r message=FALSE, warning=FALSE, fig.height = 5, fig.width= 7}
# Names of rasters in Rasters directory (everything before extension)
rasters_names_list <- list("dem",  # elevation
                          "ptoc", # proxmity to coast
                          "east", # easting
                          "north",# northing
                          "asp",  # aspect
                           "tpi", # topographic position index
                          "slope") # slope
# Make into rasters objects list
rasters_list <- lapply(FUN = raster,
                       X   = paste0("E:\\Packages\\swnsmodelr\\Rasters\\200\\",rasters_names_list, ".tif"))

# Plot brick of rasters
rasters_list %>% brick() %>% plot()


```

  The constant raster values are added to the `swns_stations_df` table with  `extract_constant_rasters_values()`.  A new frame will be made called `swns_stations_df_200`, which is the original table plus constant raster values at 200m resolution
```{r message=FALSE, warning=FALSE}
swns_stations_df_200 <- extract_constant_raster_values(swns_stations_df, rasters_list)
str(swns_stations_df_200)
```


### Temporally Changing Rasters

  In this modelling procedure, a solar radiation raster for each day is used an input into the model. The large quantity of rasters makes it problematic to store rasters as a list of raster objects, as with the constant rasters. To overcome this issue, there are two steps to adding temporally changing rasters to `swns_stations_df`. 
  In the first step, the file path to each raster is stored as a row in a new dataframe. The date is then parsed from the file path name, to a new column of corresponding dates. This way, rasters can be selected and converted to rasters objects based on dates. The data frame, referred to as a 'temporal rasters data frame', is made with `make_temporal_rasters_df()`. The user must define the "date characters" or `date_chars` argrument, which is the index of characters that contain the date in each file path. The date format (`date_format` agrument) must also be defined (see ?as.Date). The user also defines a start and end date of rasters to include, and the extension of the rasters. 
  Using the file "Sum_Irradiance_2012_1.tif" as an example: the date is formed with the characters 16 to -5, and the format is "%Y_%j". A sample use of the function for the solar radiation rasters is shown below:
  
```{r message=FALSE, warning=FALSE}
solar_irradiance_rasters_df <- make_temporal_raster_df(
  in_folder = "E:\\GOES_200m",
  start_date = ymd('2012-01-01'),
  end_date   = ymd('2012-12-31'),
  date_chars = c(16,-5),
  date_format = "%Y_%j",
  extension = ".tif")

head(solar_irradiance_rasters_df)
```

Finally, to add the values of each temporal raster to the weather station data, use `extract_temporal_raster_values()`. This may take a while if there are many temporal rasters. The function add a column for the set of temporally changing rasters. The argument `col_name` is used to specify the name of the column. By setting `verbose` to true, the function will print a completion report for every raster that has been succesfully extracted to the weather stations data frame. The output data frame was named `swns_stations_df_200`, because it has the raster values extracted from 200m resolution.

```{r eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

swns_stations_df_200  <- extract_temporal_raster_values(temporal_rasters_df = solar_irradiance_rasters_df,
                               temperatures_df = swns_stations_df_200,
                               col_name = "sum_irradiance",
                               verbose = FALSE)
```


```{r}
head(swns_stations_df_200)
```



## Separate Stations into Modelling and Validating Stations
 There are weather stations from AGRG (study stations), ECAN and DNR (external stations) in the data frame. Ideally, the study stations would be used to model the data, and the external stations to validate the models.   To be able to differeniate at any time between study (AGRG) and external (ECAN and DNR) stations, the station I.D.'s were sorted into two lists: `study_stations_list` and `ext_stations_list`. In the code below, the lists were used to separate the stations, and they were plotted to show their locations. 

```{r fig.hold=TRUE, message=FALSE, warning=FALSE,height = 7, fig.width= 7}
ggplot(data = swns_stations_df %>%
         filter(!duplicated(stationid)) %>%
         dplyr::mutate(origin = if_else(stationid %in% study_stations_list,
                                                               "study",
                                                               "ext"))) +
  geom_point(aes(x = EASTING, y = NORTHING, colour = origin)) +
  coord_fixed()

```
  

The study stations alone do not provide sufficient coverage in the eastern most portion surrounding the Halifax area, and the Annapolis valley. There are external stations in those areas which were included in the model stations along with the study stations. The external stations that were included as model stations were '6354' in Greenwood, '6456' north is St. Margaret's Bay, '47187' in Halifax, 'DN025' south of Digby and '6497' near Berwick. A  number of study stations were used in validating the models rather than generating them, these stations were: "YA4","S160","AN3","CL4","KE5","WE5","WE2","LI2","CH4". These stations are located in areas represented by model stations but not so close that their modelling values were essentially the same.


The following code was used to separate the stations, and can be easily altered to exclude or include stations in either category:
```{r}
# Choose modelling stations
model_stations_df <- swns_stations_df_200 %>%
          # 1. Include all study stations
  filter(stationid %in% study_stations_list |
          # 2. Included these ext stations
          stationid %in% c("6354","47187","6456","DNR025","6497")) %>% 
          # 3. Remove these study stations
  filter((stationid %in% c("YA4","S160","AN3","CL4","KE5","WE5","WE2","LI2","CH4")) == FALSE)

# Choose validation stations
val_stations_df   <- swns_stations_df_200 %>% 
          # 1. Include all ext stations
  filter(stationid %in% ext_stations_list  |    
          # 2. Add these study stations
           stationid %in% c("YA4","S160","AN3","CL4","KE5","WE5","WE2","LI2","CH4")) %>%
          # 3. Remove these ext stations
  filter(!(stationid %in% c("6354", "47187", "6456","27141", "DLG006", "DNR024",
                            "6501", "DNR025", "DNR003", "6497")             
           )
         ) 
```

This is the resulting modelling and validation stations:

```{r fig.hold=TRUE, message=FALSE, warning=FALSE,height = 7, fig.width= 7}
ggplot(data = swns_stations_df %>%
         filter(!duplicated(stationid)) %>%
         dplyr::mutate(role = if_else(stationid %in% unique(model_stations_df$stationid),
                                                               "model",
                                                               "validate"))) +
  geom_point(aes(x = EASTING, y = NORTHING, colour = role)) +
  coord_fixed()

```

## Examine Raster Variable Relationship with Weather Station Temperature


### Weather Station Coverage of Raster Variability
```{r, message=FALSE, warning=FALSE, fig.height = 5, fig.width= 7}
model_stations_sp <- model_stations_df %>% 
  filter(!duplicated(stationid)) %>%
  dplyr::select(stationid, EASTING, NORTHING)
coordinates(model_stations_sp) = ~ EASTING + NORTHING

par(mfrow = c(2,2))
for(i in seq_along(rasters_list)){
  raster_values <- raster::extract(rasters_list[[i]], model_stations_sp) 
  if(names(rasters_list[[i]]) == "slope"){
    plot(x = 1:(length(raster_values)),
       y = raster_values[sort.list(raster_values)], 
       ylab = names(rasters_list[[i]]),
       xlab = "Stations",
       ylim = c(0,90))
  }else{
  plot(x = 1:(length(raster_values)),
       y = raster_values[sort.list(raster_values)], 
       ylab = names(rasters_list[[i]]),
       xlab = "Stations",
       ylim = c(minValue(rasters_list[[i]]), maxValue(rasters_list[[i]])))
}
  
}
```


### Test Dependent Variable/Independent Variable Combinations at Different Timeframes

In the `swns_stations_df_200` data frame there are three dependent variables: daily minimum, maxmimum, and mean temperatures, and seven independent variables extracted from rasters. Functions were written to model dependent and independent variables one-on-one, to see the individual relationships. A function was written for the daily, weekly, and monthly timeframes: `extract_daily_spvs()`, `extract_weekly_spvs`, `extract_monthly_spvs`. The term "spvs" refers to "smooth term p-values". 
The functions a data frame that is similar to the input data frame, with three added columns: `ind_var` and `dep_var`, which specify which columns in the row were used as independent and dependent variables, respectively; and `pval`, which is the p-value from the smooth term in the GAM. 

A sample of using the three functions, and plotting their results, are in the code examples below.

```{r message=FALSE, warning=FALSE}
swns_stations_df_200 <- add_date_columns(swns_stations_df_200)
daily_spvs_df <- extract_daily_spvs(swns_stations_df_200, # data frame
                                    c("temp_min","temp_max",
                                      "temp_mean"),          # dependent variables
                                    c("dem","asp","slope","tpi","ptoc",
                                      "sum_irradiance","east, north"), #independent variables
                                    2012, # year
                                    1:10) # days in the year (ex: 1:365)
ggplot(data = daily_spvs_df, aes(x = yday, y = pval)) +
  geom_smooth(aes(colour = ind_var)) +
  facet_wrap(~dep_var, ncol = 1) +
  labs(x = "Day of the Year - 2012", y = "P-Value",
       title = "Daily Timeframe",
       colour = "Independent \n Variable")
```

```{r message=FALSE, warning=FALSE}
swns_stations_df_200 <- add_date_columns(swns_stations_df_200)
daily_spvs_df <- extract_weekly_spvs(swns_stations_df_200, # data frame
                                    c("temp_min","temp_max",
                                      "temp_mean"),          # dependent variables
                                    c("dem","asp","slope","tpi","ptoc",
                                      "sum_irradiance","east, north"), #independent variables
                                    2012, # year
                                    1:10) # weeks in the year (ex: 1:52)
ggplot(data = daily_spvs_df, aes(x = yday, y = pval)) +
  geom_smooth(aes(colour = ind_var)) +
  facet_wrap(~dep_var, ncol = 1) +
  labs(x = "Day of the Year - 2012", y = "P-Value",
       title = "Weekly Timeframe",
       colour = "Independent \n Variable")
```


```{r, eval = FALSE, message=FALSE, warning=FALSE}
swns_stations_df_200 <- add_date_columns(swns_stations_df_200)
daily_spvs_df <- extract_monthly_spvs(swns_stations_df_200, # data frame
                                    c("temp_min","temp_max",
                                      "temp_mean"),          # dependent variables
                                    c("dem","asp","tpi","ptoc",
                                      "sum_irradiance","east, north"), #independent variables
                                    2012, # year
                                    1) # months in the year (ex: 1:365)
ggplot(data = daily_spvs_df, aes(x = yday, y = pval)) +
  geom_smooth(aes(colour = ind_var)) +
  facet_wrap(~dep_var, ncol = 1) +
  labs(x = "Day of the Year - 2012", y = "P-Value",
       title = "Monthly Timeframe",
       colour = "Independent \n Variable")
```

## Generate and Validate Models 

A set of functions in `swnsmodelr` generate GAMs based on a model station data, then apply the models to validation station data, and computes the error at each validation record. The output from the functions is the validation station data frame with columns for the error (`resid`), GCV score(`gcv`), and adjusted R square () added. There is a function written for the three timeframes, daily, weekly, and monthly, respectively: `validate_daily_GAMs()`, `validate_weekly_GAMs()`, `validate_monthly_GAMs()`.

The functions allow the user to specify an alternative formula, with the argument `alt_formula`. This feature was specifically implemented for the case that solar radiation data is missing for a particular date. The function will fail if the data is missing, and fall back on the alternative formula. If there is no suspiscion that a missing variable will cause the function to fail, this argument can be left blank. 

```{r echo=TRUE}
daily_val_df <- validate_daily_GAMs(model_stations_df,
                                     val_stations_df,
                                     years = 2012,
                                     days = 1:3,
                                    formula = "temp_mean ~ s(dem) + s(asp) + s(tpi) 
                                    + s(ptoc, k= 3) + s(east,north) + s(sum_irradiance)",
                                    alt_formula ="temp_mean ~ s(dem) + s(asp) + s(tpi) 
                                    + s(ptoc, k = 3) + s(east,north)")
```


Two ways of visualizing the errors are by histogram, and a scatterplot against the dependent variable.
```{r message=FALSE, warning=FALSE}
ggplot(data = daily_val_df, aes(resid))+
  labs(x = "Residuals",y = "Count") +
  geom_histogram(position = "dodge", colour = "black", fill = "pink",bins = 30)

ggplot(data = daily_val_df, aes(x = temp_mean, y = resid)) +
  geom_point() +
  geom_smooth()
```


## Generating Output
### Overview
With the combination of a model stations data frame, model formula, and input rasters, functions in `swnsmodelr` generate different types of raster outputs.  The functions `daily_model_predict_rasters()`, and `weekly_model_predict_rasters()` generates daily rasters of the dependent variable with the specified model applied in daily or weekly timeframes, respectively. 

The prediction functions will fail if the rasters in the `rasters_list`, and the rasters pointed to in the `temporal_rasters_df`, do not have the same extents, and number of rows and columns. The function `raster::compareRaster()` will be return `TRUE` for the `rasters_list` if the properties are the same. If not, the function `raster::resample()` can be used to correct the issue. 

The function `generate_mean_rasters()`, calculates mean rasters for sets of maximum and minimum rasters that correpsond by date. This function was written incase a user would rather model daily minimum and maximum temperatures before calculating daily mean temperature. 

Finally, the function `generate_output_gdd()`, generates GDD rasters from daily mean temperature rasters. The user can specify the GDD base, and frequency of outputs (daily, weekly, monthly or yearly). There is also an agrument called `growing_season` that when set to `TRUE`, only accumulates GDD for April to November.  

### Predict Daily Mean Temperature Rasters
Continuing from the previous section, the code below demonstrates how to use `daily_model_predict_rasters()`.As with the model validation, a formula and an alternative formula (`alt_formula`) can be specified. The `alt_formula` is used if the `formula` causes `gam()` to fail. This would occur is a variable is a missing for the date being modelling, which is the case randomly with `sum_irradiance`. 

```{r eval = FALSE}

# Predict daily mean temperature
daily_model_predict_rasters(start_date = ymd('2012-01-01'),
                               end_date = ymd('2017-12-31'),
                               formula = "temp_mean ~
                                          s(east,north) +
                                          s(dem) +
                                          s(sum_irradiance) +
                                          s(ptoc, k= 3)",
                               alt_formula = "temp_mean ~
                                          s(east,north) +
                                          s(dem) +
                                          s(ptoc, k= 3)",
                               temperatures_df = model_stations_df,
                               input_rasters = rasters_list,
                               temporal_rasters_df = solar_irradiance_rasters_df,
                               output_folder = "F://daily_models//mean_temperature",
                               output_ext = "tif",
                               verbose = TRUE)
```


### Calculate Growing Degree Day (GDD) Rasters

Rasters of GDD are calculated by raster algebra with the daily mean temperature rasters. The first step is to make a data frame with `make_temporal_rasters_df()` that contains the file paths to the daily mean temperature rasters, and their corresponding date.
The function `output_gdd_rasters()` applies the specified GDD base to the mean temperature rasters, accumulates them daily, and sends them to the output folder at the specified timeframe, `output_timeframe`. The accumulation only occurs within a year, and can be set to only include April 1st - November 30th by setting `growing_season` to `TRUE`. The code below demonstrates how to use the two functions to generate GDD rasters.

```{r eval = FALSE}
temp_mean_df <- make_temporal_raster_df("F://daily_models//mean_temperature",
                                        start_date = ymd('2012-01-01'),
                                        end_date   = ymd('2017-12-31'),
                                        date_chars = c(10,19),
                                        date_format = "%Y-%m-%d")

generate_gdd_output(temp_mean_df,
                    gdd_base = 5,
                    start_date = start_date,
                    end_date = end_date,
                    output_time_slice = "monthly",
                    growing_season = TRUE,
                    output_folder = "Z:\\Dana\\Weekly\\Monthly_GDD_200_7",
                    plot_gdd_raster = TRUE)
```


