---
title: "Selecting and Importing Environmental Rasters for Modelling"
output: html_notebook
---
```{r message=FALSE, warning=FALSE}
library(swnsmodelr)
```

Temperature is related to several features in the environment which already have raster representations. These rasters can be used to predict temperature at unknown locations. The values of the rasters cells coincident with weather stations are used to generate models, then the model are carried over the remainder of the raster to interpolate a temperature surface.
Here, a distinction is made between rasters that represent a constant environmental feature over time,
and those that are changing through time.

## Constant Rasters
Rasters of environmental features that remanin constant throughout time are referred to as "constant rasters". The constant rasters that are used in modelling temperature here are: elevation, aspect, easting\*, northing\*, and proximity to coast. (\* easting and northing values in UTM zone 20 N)

The constant rasters are available in the SWNS_Modelling project, and can be brought into the RStudio environment using the code below. Note that the rasters are imported together as a list of raster objects. To add a raster to the list from another location, either first add the raster to the Rasters directory in the project and include it in the names list, or create the raster object as the next available element in the list.
```{r message=FALSE, warning=FALSE}
# Names of rasters in Rasters directory (everything before extension)
rasters_names_list <- list("dem",  # elevation
                          "ptoc", # proxmity to coast
                          "east", # easting
                          "north",# northing
                          "asp",  # aspect
                           "tpi",
                          "slope")
rasters_list <- lapply(FUN = raster,
                       X   = paste0("F:\\Packages\\swnsmodelr\\Rasters\\200\\",rasters_names_list, ".tif"))

rasters_list %>% brick() %>% plot()


```



### Examining Relationships Between Temperature and Constant Rasters

Selecting environemental rasters that will help in modelling temperature is based on intuition and exploratory data analysis. The rasters need to both: have a wide range of their values represented at weather station locations; and have some kind of mathematical relationship between their values and those at weather stations.

#### Are a wide range of raster values represented at weather stations?
For each raster, the values at the weather stations will be extracted, ordered, and plotted so we can get a sense of the covered range.
```{r message=FALSE, warning=FALSE}
par(mfrow = c(2,2))
for(i in seq_along(rasters_list)){
  raster_values <- raster::extract(rasters_list[[i]], model_stations_sp) 
  plot(x = 1:(length(raster_values)),
       y = raster_values[sort.list(raster_values)], 
       ylab = names(rasters_list[[i]]),
       xlab = "Stations")

  
}
```


#### Do the raster values at weather stations correlate with temperature?

Using a function from the swnsmodelr package, the values from a set of constant rasters can be easily attached to the dataframe of temperature records. 

```{r message=FALSE, warning=FALSE}
df <-extract_constant_raster_values(daily_temperatures_df, rasters_list)
df <- add_date_columns(df)
head(df)
```




Scatter plots
```{r message=FALSE, warning=FALSE}

df_week <- df %>% filter(week == 26,
                         year == 2012)
df_week_melt <- melt(id = c("stationid", "date_time", "temp_mean","temp_min","temp_max","EASTING","NORTHING",
                            "yday","week","month","year"),
                     data = df_week, na.rm = TRUE)

ggplot(data = df_week_melt, mapping = aes(x = value, y = temp_mean, colour = yday)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~variable, scales = "free")
    

```



```{r fig.height=10, message=FALSE, warning=FALSE}

df_week <- df %>% filter(week == 20,
                            year == 2012)

gam_m <- gam(temp_mean ~
               s(dem, yday) +
               s(tpi, yday) +
               s(east,north) +
               s(asp_c, yday) +
               s(ptoc),
             data = df_week)
par(mfrow = c(2,3))
plot(gam_m)
summary(gam_m)


```




## Temporal Rasters

Other conditions of the environment are changing on a daily basis, and therefore have a differet raster for each day. In this project, solar irradiance is a set of rasters that change over time, or temporal rasters. There is a solar irradiance raster for each day, therefore there is a large demand for memory for the entire set of rasters. To negate this obstacle, the set of temporal rasters will be stored as a data frame of path names and dates. This way, the rasters remain simply as a path string, with an associated date, until it need be converted to a raster object. 


The function 'make_temporal_raster_df' makes a dataframe of path names and dates from a folder of temporal rasters.

```{r}
solar_irradiance_rasters_df <- make_temporal_raster_df(
  in_folder = "F:\\GOES_200",
  start_date = ymd('2012-01-01'),
  end_date   = ymd('2017-12-31'),
  date_chars = c(16,-5),
  date_format = "%Y_%j",
  extension = ".tif")

head(solar_irradiance_rasters_df)
```

The key to the function is the user supplied parsing method for the date. The argument "date_chars" tells the function the first and late characters that make up the date component in the file name. In combination with the self-explantory "date_format" argument, a date can be parsed from a file name and converted to a date class variable. The code below demonstrates how to figure out the date_chars and test it.

```{r}
temporal_rasters_folder <- "F:\\GOES_200"
test_file_name <- list.files(temporal_rasters_folder)[[1]]

date_chars <- c(15, -4)
date_format <- "%Y-%j"  # year and day of year separated by "-"

test_file_name %>% str_sub(date_chars[[1]], date_chars[[2]]) %>%
                   print()


test_file_name %>% str_sub(date_chars[[1]], date_chars[[2]]) %>%
                   as.Date(format = date_format) %>%
                   print()

```

From the above print out, clearly the date_chars is off. By adjusting them, the correct characters can be subsetted from the file name.



```{r}
temporal_rasters_folder <- "F:\\GOES_200"
test_file_name <- list.files(temporal_rasters_folder)[[1]]

date_chars <- c(16, -5)
date_format <- "%Y-%j"  # year and day of year separated by "-"

test_file_name %>% str_sub(date_chars[[1]], date_chars[[2]]) %>%
                   print()

test_file_name %>% str_sub(date_chars[[1]], date_chars[[2]]) %>%
                   as.Date(format = date_format) %>%
                   print()
```



This time, the date_format is wrong so the second print out is NA. The date format shows the components separated by a hyphen, when in the file name, it is actually an underscore.
```{r}
temporal_rasters_folder <- "F:\\GOES_200"
test_file_name <- list.files(temporal_rasters_folder)[[1]]

date_chars <- c(16, -5)
date_format <- "%Y_%j"  # year and day of year separated by "-"

test_file_name %>% str_sub(date_chars[[1]], date_chars[[2]]) %>%
                   print()

test_file_name %>% str_sub(date_chars[[1]], date_chars[[2]]) %>%
                   as.Date(format = date_format) %>%
                   print()
```



The function extract_temporal_raster_values uses the output of make_temporal_rasters_df and a SpatialPointsDataFrame

```{r eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE}

df   <- extract_temporal_raster_values(temporal_rasters_df = solar_irradiance_rasters_df,
                               temperatures_df = df,
                               col_name = "sum_irradiance",
                               stations = stations_sp,
                               verbose = FALSE)
```


Make plots to see how solar irradiance and temp_mean correlates

```{r message=FALSE, warning=FALSE}
df <- add_date_columns(df)
df_subset <- df %>% filter(week == 14,
                           sum_irradiance != 0 & is.na(sum_irradiance) == FALSE) %>%
                mutate(as.character(yday))
ggplot(data = df_subset, mapping = aes(x = sum_irradiance, y = temp_mean)) +
  geom_point() +
  geom_smooth(method = gam) +
  facet_wrap(~yday, scale = "free")
```

